---
title: "Assignment 1 ACST886"
author: "Vu Nguyen"
date: "21/08/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

I apply the annuity 
$R\times\sum_{n = 1}^{30} v^n=600,000$
In which:  $R$ is the annual payment  
  $v$ is the discount factor with $v=(1+i)^{-1}$
  
```{r}
pmt=function(R) {
  i=0.035
  v=1/(1+i)
  n=c(1:30)
  a=sum(v^n)
  R*a-600000
}
```
The code create a function simulating the equation: 
$R\times\sum_{n = 1}^{30} v^n-600,000=0$  

Now we use uniroot command to find the solution for the equation.
```{r}
Q1=uniroot(pmt,interval = c(1,9999999))$root
Q1
```
The part $root extract the solution from the result table.
Testing the result:
```{r}
pmt(Q1)
```
This is the correct result.

# Question 2
## Part a)
Since Scott take on the same loan as Bill in Question 1 with the same term and interest rate, by not including the interest offset amount the initial payment of Scott and Bill should be the same.
```{r}
Q1
```
## Part b)
Since Scott have an offset account which pay annual interest at 3.5%, this means the annual repayment of Scott would be $pmt+100,000\times3.5\%=pmt+3,500$. With a greater annual repayment, the payment term should be less than 30 years. Prove:
```{r}
nper=function(n) {
  i=0.035
  v=1/(1+i)
    a=(1-v^n)/i
  (Q1+3500)*a-600000
}
```
Since we expecting a non-integer anwser, I switch to an annuity formulae.
The code create a function simulating the equation: 
$(R+3,500)\times\frac{1-v^n}{i}-600,000=0$  
Note that: $=\frac{1-v^n}{i}=\sum_{j = 1}^{n}v^j$

```{r}
Q2a=uniroot(nper,interval = c(1,9999999))$root
Q2a
ceiling(Q2a)
```
The ceiling command round up the answer since the term can only be an interger. As we can see it only takes 26 years for Scott to payoff the loan  

Since the calculated payment term is a non-integer, the last payment of Scott should be less than the initial payment. Prove:

```{r}
lpmt=function(z) {
  i=0.035
  v=1/(1+i)
    a=(1-v^25)/i
  (600000-(Q1+3500)*a)*(1+i)^26-z
}
```

I use the restrospective method to calculate the loan outstanding right before the last payment.
The code create a function simulating the equation: 
$[[600,000-(R+3,500)]\times\frac{1-v^{25}}{i}]\times(1+i)^{26}-z=0$
In which: $z$ is the last repayment including the offset amount.

```{r}
Q2b=uniroot(lpmt,interval = c(1,9999999))$root
Q2b
```
The result shows that the last payment is less than that calculated in part a).
In conclusion, with the offset account in effect, the payment term of Scott would be 26 years with the last repayment is $11,353.07

# Question 3
## Part a)
Theresa receives $1,000 at the end of each year for 30 years. The total present value of the commission would be:  
$1,000\times\frac{1-v^{30}}{i}$  
In which:  
$i=3.5\%$ and $v=(1+i)^{-1}$
```{r}
i=0.035
v=1/(1+i)
Q3a=1000*(1-v^30)/i
Q3a
```

## Part b)
We have:
$L=R\times\sum_{n = 1}^{n} (1+i)^{-j}$  
In which:  
$L$ is the total loan amount  
$n$ is the loan term  
$i$ is the rate  
$R$ is the periodical repayment amount  
As you can see, as the loan amount $L$ increase and term $n$ and $R$ stay the same, the effective interest $i$ will decrease

## Part c)
To calculate the effective rate, we will create a function simulate the equation on part b).
```{r}
rate=function(i) {
  v=1/(1+i)
  n=c(1:30)
  a=sum(v^n)
  Q1*a-600000-Q3a
}
```
The code create a function simulating the equation:
$R\times\sum_{n = 1}^{n} (1+i)^{-j}-L=0$  
In which:  
$L$ is the sum of the loan amount and the total PV of Theresa commission at 19 March 2019.
```{r}
Q3b=uniroot(rate,interval = c(0.000000001,9999999))$root
Q3b
```
The uniroot command  will find a interest rate $i$ which with all the given input ($L, n, R$), the equation simulated in function $rate$ will result in 0. This will be our effective rate.

## Part c)
If Bill was offer the interest rate in part b), his annual repayment (assuming the loan term and amount stay the same) would be:
```{r}
i=Q3b
v=1/(1+i)
n=c(1:30)
Q3c=600000/sum(v^n)
Q3c
```
The new annual repayment is $32,622.8
```{r}
(Q1-Q3c)/Q1
```
Bill's annual repayment decrease by 2.97%

# Question 4
Input data
```{r}
LifeA <- as.Date(c("1965-05-16", "1985-07-21", "1994-06-22"))
LifeB <- as.Date(c("1922-03-01", "1955-09-23", "1993-07-21"))
LifeC <- as.Date(c("1942-09-21", "1967-06-02", "1996-04-29"))
```
## Part a) Age next birthday at death
```{r}
LifePartA=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-montha)*32+(dayc-daya))>=0){
  yearc-yeara+1
} else { yearc-yeara}
}
```
- The first line is to create a function to calculate age next birthday
- The next 6 lines are to extract the date, month, year value for calculate.
- The 8th line indicate the condition whether the Life die before or after birthday within the year of dead. The next 2 lines give coresponding results.
The code to check the result is as below.
```{r}
data.frame("Age of life A"=LifePartA(LifeA), "Age of life B"=LifePartA(LifeB), "Age of life C"=LifePartA(LifeC))
```


## Part b) Age last birthday an 1 January preceding death
```{r}
LifePartB=function(Life){
montha=as.numeric(substr(Life[1],6,7))
daya=as.numeric(substr(Life[1],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (montha!=1&&daya!=1){
  yearc-yeara-1
} else { yearc-yeara}
}
```
- I extract the value just like in part a)  
- For the condition, if the birthday of the life chosen is not in 01 January, we simple take the difference between death year and birth year minus 1. Otherwise, we won't minus one

```{r}
data.frame("Age of life A"=LifePartB(LifeA), "Age of life B"=LifePartB(LifeB), "Age of life C"=LifePartB(LifeC))

```

## Part c) Age at the birthday in the policy year of death
```{r}
LifePartC=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthb=as.numeric(substr(Life[2],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayb=as.numeric(substr(Life[2],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearb=as.numeric(substr(Life[2],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-monthb)*32+(dayc-dayb))>0){
  z=yearc
} else {z=yearc-1}
if (((monthb-montha)*32+(dayb-daya))>0){
  z+1-yeara
} else {z-yeara}
}

```
-The first 10 line is to create the function to calculate age and extract the date value like part a and part b.

```{r}
data.frame("Age of life A"=LifePartC(LifeA), "Age of life B"=LifePartC(LifeB), "Age of life C"=LifePartC(LifeC))
```

