---
title: "Assignment 2_ACST886"
author: "Vu Nguyen"
date: "04/09/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is the link to my Assignment on Github:  
https://github.com/vunguyen1502/ACST886/tree/master/Assignment%202  
https://github.com/vunguyen1502/ACST886  
Either link would do.  

# Question 1
First we input the data.  
```{r}
LifeA <- as.Date(c("1964-03-17", "1992-06-20", "1998-06-20"))
LifeB <- as.Date(c("1964-05-06", "1992-08-06", "1993-06-12"))
LifeC <- as.Date(c("1964-08-12", "1992-12-18", "1995-06-18"))
LifeD <- as.Date(c("1964-10-27", "1993-01-04", "1998-06-20"))
LifeE <- as.Date(c("1965-01-04", "1993-04-28", "1996-08-29"))
LifeF <- as.Date(c("1965-04-18", "1993-06-16", "1995-12-12"))
LifeG <- as.Date(c("1965-05-20", "1993-10-29", "1996-04-21"))
LifeH <- as.Date(c("1965-07-04", "1994-02-16", "1998-06-20"))
LifeI <- as.Date(c("1965-09-16", "1994-08-22", "1997-02-22"))
LifeJ <- as.Date(c("1965-12-11", "1995-03-06", "1997-02-17"))
```
I add the exit day for those who live on without withdraw which is the anniversaries in 1998.
We will need to calculate the initial exposure for each life.  

## Calculate eposure for those who live without withdrawing
I create a function to calculate the exposure for those who live on without withdrawing.
```{r}
live=function(Life){
  insure=as.Date(paste0(substr(Life[1],1,4),substr(Life[2],5,10)))
  if (abs(as.numeric(Life[1]-insure))<183){
  insbirth=as.numeric(substr(Life[1],1,4))
  } else if (as.numeric(Life[1]-insure)>182){
  insbirth=as.numeric(substr(Life[1],1,4))+1
  } else if (as.numeric(Life[1]-insure)<-182){
  insbirth=as.numeric(substr(Life[1],1,4))-1
  }
  agexentry=as.Date(paste0(insbirth+30,substr(Life[2],5,10)))
  min(as.numeric(Life[3]-agexentry)/365.25,1)
}

```
- I create a date which is the policy anniversary at birth year. (insure)
- The next 6 line is the condition to find the birth year where Life's isuring age is exactly 0. Basically, the condition find the anniversary date nearest to the birthday. 
- agexentry is the policy anniversary where age is exactly 30.  
- The last line is to take the date of exit (anniversay at 1998) minus the age 30 date of entry (agexentry). If it is greater than 1, then they have full exposure at age 30.  
$~$
Now we calculate the exposure for those who live without withdrawing.  
```{r}
AgeLive=(live(LifeA)+live(LifeD)+live(LifeH))
AgeLive
```

## Calculate eposure for those who withdrawed
Now, I create a function to calculate the exposure for those who withdrawed.
```{r}
withdrawal=function(Life){
  insure=as.Date(paste0(substr(Life[1],1,4),substr(Life[2],5,10)))
  if (abs(as.numeric(Life[1]-insure))<183){
    insbirth=as.numeric(substr(Life[1],1,4))
  } else if (as.numeric(Life[1]-insure)>182){
    insbirth=as.numeric(substr(Life[1],1,4))+1
  } else if (as.numeric(Life[1]-insure)<-182){
    insbirth=as.numeric(substr(Life[1],1,4))-1
  }
  policyyear=as.numeric(substr(Life[2],1,4))
  issuringage=policyyear-insbirth
  withdrawage=as.numeric(substr(Life[3],1,4))-policyyear+issuringage
  agexentry=as.Date(paste0(insbirth+30,substr(Life[2],5,10)))
  agexexit=as.Date(paste0(insbirth+withdrawage,substr(Life[2],5,10)))
  min(max(as.numeric(agexexit-agexentry)/365.25,0),1)
}
```
- I create a date which is the policy anniversary at birth year. (insure)
- The next 6 line is the condition to find the birth year where Life's isuring age is exactly 0. Basically, the condition find the anniversary date nearest to the birthday.
- The next 2 lines are to caculate the sd age (issuringage)
- The I calculate the age a withdrawal (withdrawage)
- agexentry is the policy anniversary where age is exactly 30.
- agexexit is the policy anniversary where Life withdraw from the policy.
- The last line is to take the date of exit (anniversay at withdrawal) minus the age 30 date of entry (agexentry). If it is greater than 1, then they have full exposure at age 30. Otherwise, divide it for 365.25 to get eposure time.
$~$  
$~$  
$~$  
$~$
Now we calculate the exposure for those who withdrawed.
```{r}
AgeWithdraw=(withdrawal(LifeC)+withdrawal(LifeF)+withdrawal(LifeI))
AgeWithdraw
```
## Calculate eposure for those who died
I create a function to calculate the exposure for those who died.  
```{r}
death=function(Life){
  insure=as.Date(paste0(substr(Life[1],1,4),substr(Life[2],5,10)))
  if (abs(as.numeric(Life[1]-insure))<183){
    insbirth=as.numeric(substr(Life[1],1,4))
  } else if (as.numeric(Life[1]-insure)>182){
    insbirth=as.numeric(substr(Life[1],1,4))+1
  } else if (as.numeric(Life[1]-insure)<(-182)){
    insbirth=as.numeric(substr(Life[1],1,4))-1
  }
  agexentry=as.Date(paste0(insbirth+30,substr(Life[2],5,10)))
  insuredeath=as.Date(paste0(substr(Life[3],1,4),substr(Life[2],5,10)))
  if (abs(as.numeric(Life[3]-insuredeath))>0){
    insdeath=as.numeric(substr(Life[3],1,4))+1
  } else {
    insdeath=as.numeric(substr(Life[3],1,4))
  }
  agexentry=as.Date(paste0(insbirth+30,substr(Life[2],5,10)))
  agexexit=as.Date(paste0(insuredeath,substr(Life[2],5,10)))
  min(max(as.numeric(agexexit-agexentry)/365.25,0),1)
}
```
- I create a date which is the policy anniversary at birth year. (insure).  
- The next 6 line is the condition to find the birth year where Life's isuring age is exactly 0. Basically, the condition find the anniversary date nearest to the birthday.  
- I create a date which is the policy anniversary at death year. (insuredeath).  
- The next 6 line is the condition to find the year of the policy anniversary following the death date.  
- The last line is to take the date of exit (anniversay following death) minus the age 30 date of entry (agexentry). If it is greater than 1, then they have full exposure at age 30.   
$~$
Now we calculate the exposure for those who died.
```{r}
AgeDeath=death(LifeB)+death(LifeE)+death(LifeG)+death(LifeJ)
AgeDeath
```
Now we estimate $q_{30}$ with $q_{30}=\frac{d_x}{E_x}$
```{r}
qx=2/(AgeDeath+AgeLive+AgeWithdraw)
qx
```
Since life B died before observation period (anniversary 1993 to 1998) and life J died after turning age 31, We will not be adding them to the dead count.

## Estimate q using maximum likelihood.
We round the exposure of those who age 30 in leap year to 1 for easy calculation.  
The likelihood function is:  
$L=q^2(1-q)^6$  
$\frac{dL}{dq}=2q(1-q)^6+6q^2(q-1)^5$  
Our estimate $q$ is $\hat{q}$ where $\frac{dL}{dq}=2\hat{q}(1-\hat{q})^6+6\hat{q}^2(\hat{q}-1)^5=0$  
We will create a function and use command uniroot for this.  
```{r}
maxlh=function(q){
  2*q*((1-q)^6)+(q^2)*6*((q-1)^5)}
uniroot(maxlh,c(0.01,0.99))$root
```
As we can see, the result from both method are similar which can be round as $q_{30}=25\%$

# Question 2

First, we input the data.  
```{r}
Q2=data.frame("Age-group"=c("40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80-84","85-89","90-94"),
                "Exposed to Risk"=c(15518,19428,21594,21890,19174,15775,11414,6993,3276,1096,201),
                "Actual deaths"=c(65,144,219,378,465,557,685,644,471,217,67),
                "Expected deaths"=c(73.9,134.6,223.9,346.3,468.1,600.2,675.5,637.4,458.7,240.6,61.4))
Q2
```
Make sure to set up some vector for easy coding
```{r}
Ex=Q2$Exposed.to.Risk
dx=Q2$Actual.deaths
E=Q2$Expected.deaths
qx=E/Ex
```
Now we can do some testing:  
$~$  
$~$  
$~$  

## Chi-Square Test
The chi-square test statistic is:  
$\sum_{x}^{} z_{x}$  
```{r}
z=(dx-qx*Ex)/sqrt(Ex*qx*(1-qx))
sum(z^2)
```
Now compare it with 95th percentile of chi square distribution with several different parameter
```{r}
qchisq(0.95,11)
qchisq(0.95,10)
qchisq(0.95,9)
qchisq(0.95,8)
qchisq(0.95,7)

```
As you can see, the statistic is smaller than the 95 percentile of chi square with considerable reduce in degree of freedom.  
Therefore, we retain the null hypothesis.  
However, chi-square test have several problem in the consistency of deviation size and nature. Hence, further test is required. 

## Standardised Deviations Test  

The standardised deviations test statistic:  
$\sum_{All\;interval}^{}\frac{(A-E)^2}{E}$  
First we will count the actual number falling into differnce interval
```{r}
A=c(length(z[z<(-3)]),
    length(z[z<(-2)&z>(-3)]),
    length(z[z<(-1)&z>(-2)]),
    length(z[z<0&z>(-1)]),
    length(z[z<1&z>0]),
    length(z[z<2&z>1]),
    length(z[z<3&z>2]),
    length(z[z>3]))
A
```
Now we will create the expected count
```{r}
E=11*c(0,0.02,0.14,0.34,0.34,0.14,0.02,0)
E
```
We obmit the first and last interval since there are no values in them.  
Now we calculate the statistic.
```{r}
E=E[c(2:7)]
A=A[c(2:7)]
sum(((E-A)^2)/E)
```

Now compare with the 95th quantile chi square with the degree of freedom 5 since we only use 6 interval for the statistic
```{r}
qchisq(0.95,5)
```
The statistic is smaller than the 95th quantile.
```{r}
length(z[z<(2/3)&z>(-2/3)])
length(z[z<0])
```
Testing addtional properties show that approximately half of the deviations should fall in the interval (–2/3,2/3) and the number of positive deviations  is roughly equal numbers of negative deviations.  
Both the results show evidence supporting the null hypothesis.

## Cumulative Deviations Test
$\frac{\sum_{x}^{}(d_x-E_xq_x)}{\sqrt{\sum_{x}^{}E_xq_x(1-q_x)}}$  
```{r}
sum(dx-Ex*qx)/sqrt(sum(Ex*qx*(1-qx)))
```
We can see that the statistic is smaller than 1.64 (95th quantile of standard normal). Hence, we retain the null hypothesis.

## Signs Test  
The statistic the the number of positive $z_x$ 
```{r}
length(z[z>0])
```
Now we will find the smallest value $k$ for this equation:  
$\sum_{j=0}^{k}\binom{m}{j}0.5^m\ge0.025$  
In which: $m=11$
```{r}
binomk=function(k){
  m=11
  a=c(0:k)
  fac=choose(m,a)
  (sum(fac*(0.5^m))-0.025)
}
v=numeric()
for (i in c(0:11)) {
  v[i+1]=(binomk(i))
  }
v
```
As we can see, our $k$ value is 2. We have $m-k=9$ which is greater than the statistic (6). Therefore, we retain the null hypothesis.

## Grouping of Signs Test
The statistic is simply the number of distinct groups of positive $z_x$
```{r}
z>0

```
We can see that the statistic is 4.  
Now we find the smallest $k$ value for the following equation:  
$\sum_{j=0}^{k}\binom{n_1-1}{j-1}\binom{n_2+1}{j}/\binom{n_1+n_2}{n_1}\ge0.05$  
```{r}
groupsign=function(k){
  j=c(1:k)
  n1=length(z[z>0])
  n2=length(z[z<0])
sum((choose(n1-1,j-1)*choose(n2+1,j))/choose(n1+n2,n1))-0.05
}
g=numeric()
for (i in c(1:11)) {
  g[i+1]=(groupsign(i))
}
g
```
From the result, we can see that $k=2$ which is smaller than the statistic (4). Hence, we retain the null hypothesis.

## Serial Correlations Test
The serial correlations test statistic is:  
$\frac{\sum_{i=1}^{m-1} (z_i-\bar{z}^{(1)})(z_{i+1}-\bar{z}^{(2)})}{\sqrt{\sum_{i=1}^{m-1} (z_i-\bar{z}^{(1)})^2\sum_{i=1}^{m-1} (z_{i+1}-\bar{z}^{(2)})^2}}\sqrt{m-1}$  
In which:  
$\bar{z}^{(1)}=\sum_{i=1}^{m-1}z_i/(m-1)$  
$\bar{z}^{(2)}=\sum_{i=1}^{m-1}z_{i+1}/(m-1)$  
```{r}
m=11
zi=z[c(1:10)]
zi1=z[c(2:11)]

zb1=sum(zi/(m-1))
zb2=sum(zi1/(m-1))

sqrt(m-1)*sum((zi-zb1)*(zi1-zb2))/sqrt(sum((zi-zb1)^2)*sum((zi1-zb2)^2))

```
We can see that the statistic is smaller than 1.64 (95th quantile of standard normal). Therefore, we retain the null hypothesis.
$~$

## Conclusion  

In conclusion, all test show evidence support the null hypothes which means graduated estimates are sufficiently close to the crude estimates from which they are derived, and the adherence to data is sufficient.

# Question 3
## Question 5.7  
First we input the data.  
```{r}
ax=c(70:75)
aEx=c(1000,1005,1010,1008,1006,998)
adx=c(80,90,95,105,115,125)
```
Under Gompertz’ Law, it is assumed that:
$\dot{u}_x=Bc^x$

We create a log likelihood function of the Poisson model:  
$lnL=\sum_{x=70}^{75}(-E_{(x)}^cBc^x+\theta_{(x)}ln(E_{(x)}^cBc^x)-ln(\theta_{(x)}!))$
```{r}
Q5.7=function(u){
  -sum(-aEx*u[1]*(u[2]^ax)+adx*log(aEx*u[1]*(u[2]^ax))-log(factorial(adx)))
}
```
Since we are going to use nlm to estimate maximum likelihood, I put a negative sign in front of the function so that minimizing a nagetive function will give us a maximum estimate for a positive function.  
Now we will use the command nlm to minimize the negative likelihood function to find the optimize value for $B$ and $c$

```{r}
aiv=c(0.00016,1.09)
options(warn=-1)
u=nlm(Q5.7,aiv)$estimate
u
```
aiv vector is the initial value for nlm command. 
The option command is to ignore the warning (for presentation purpose only). The nlm command will find the value B,c for vector u(B,c) so that the negative likelihood (negative minimum = positive maximum) function Q5.7 is smallest.  
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of force of mortality for each age.
```{r}
ux=u[1]*u[2]^ax
data.frame("Age"=ax,"Force of mortality"=ux)
```

## Question 5.8
As given, we have:  
$ln(\dot{q}_x/(1-\dot{q}_x)=a+bx$  
First, we input the data.
```{r}
bx=c(30:49)
bEx=c(70000,66672,68375,65420,61779,66091,68514,69560,65000,66279,67300,65368,65391,62917,66537,62302,62145,63856,61097,61110)
bdx=c(39,43,34,31,23,50,48,43,48,47,62,63,84,86,120,121,122,162,151,184)
bcrude=bdx/bEx
lnq=log(bcrude/(1-bcrude))
```

Then, we create a weighted least squares function to minimize.  
$\sum_{x=30}^{49} E_x[ln(\hat{q_x}/1-\hat{q_x})-ln(\dot{q_x}/1-\dot{q_x})]^2$  
$\sum_{x=30}^{49} E_x[ln(\hat{q_x}/1-\hat{q_x})-a-bx]^2$  
In which: $\hat{q}=d_x/E_x$
```{r}
Q5.8=function(z){
  c=z[1]+z[2]*bx
  sum(bEx*(lnq-c)^2)
}
```

Now we will use the command nlm to minimize the weighted least square function to find the optimize value for $a$ and $b$
```{r}
biv=c(-10,0.09)
z=nlm(Q5.8,biv)$estimate
z
```
biv vector is the initial value for nlm command.  
The nlm command will find the value a,b for vector z(a,b) so that the function weighted least square Q5.8 will be minimize.
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of mortality rate for each age.

We have
$ln(\dot{q}_x/(1-\dot{q}_x))=a+bx$  
Therefore,
$\dot{q}_x=e^{a+bx}/(1+e^{a+bx})$
```{r}
e=exp(z[1]+z[2]*bx)
bqx=e/(e+1)
data.frame("Age"=bx,"Mortality rate"=bqx)
```
## Question 5.9
As given, we have:  
$\dot{q}_x=a+bq_x^S$  
First, we input the data.
```{r}
cx=c(47:67)
cEx=c(166,187,218,243,276,302,347,390,430,494,558,628,701,813,917,1040,1182,1299,1432,1596,1752)
cdx=c(2,2,4,6,2,4,7,3,9,9,8,11,14,18,18,24,30,43,41,54,64)
cqx=cdx/cEx
cqsx=c(0.00505,0.00570,0.00644,0.00728,0.00826,0.00930,0.01051,0.01184,0.01331,0.01492,0.01668,0.01859,0.02065,0.02287,0.02525,0.02778,0.03049,0.03339,0.03648,0.03978,0.04332)

```
Then, we create a weighted least squares function to minimize.  
$\sum_{x=47}^{67} E_x(\hat{q_x}-\dot{q_x})^2$  
$=\sum_{x=47}^{67} E_x(\hat{q_x}-a-bq_x^S)^2$  
In which: $\hat{q}=d_x/E_x$
```{r}
Q5.9=function(r){
  qs=r[1]+r[2]*cqsx
  sum(cEx*(cqx-qs)^2)
}
```
Now we will use the command nlm to minimize the weighted least square function to find the optimize value for $a$ and $b$
```{r}
civ=c(0.007,0.6)
r=nlm(Q5.9,civ)$estimate
r
```

civ vector is the initial value for nlm command.  
The nlm command will find the value a,b for vector z(a,b) so that the function weighted least square Q5.9 will be minimize.
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of mortality rate for each age.

We have
$\dot{q}_x=a+bq_x^S$  
```{r}
qs=r[1]+r[2]*cqsx
data.frame("Age"=cx,"Mortality rate"=qs)
```

