---
title: "Question 3 Assignment 2_ACST886"
author: "Vu Nguyen"
date: "05/09/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the link to my Assignment on Github:  
https://github.com/vunguyen1502/ACST886/tree/master/Assignment%202  
https://github.com/vunguyen1502/ACST886  
Either link would do.  

# Question 3
## Question 5.7  
First we input the data.  
```{r}
ax=c(70:75)
aEx=c(1000,1005,1010,1008,1006,998)
adx=c(80,90,95,105,115,125)
```
Under Gompertz Law, it is assumed that:
$\dot{u}_x=Bc^x$

We create a log likelihood function of the Poisson model:  
$lnL=\sum_{x=70}^{75}(-E_{(x)}^cBc^x+\theta_{(x)}ln(E_{(x)}^cBc^x)-ln(\theta_{(x)}!))$
```{r}
Q5.7=function(u){
  -sum(-aEx*u[1]*(u[2]^ax)+adx*log(aEx*u[1]*(u[2]^ax))-log(factorial(adx)))
}
```
Since we are going to use nlm to estimate maximum likelihood, I put a negative sign in front of the function so that minimizing a nagetive function will give us a maximum estimate for a positive function.  
Now we will use the command nlm to minimize the negative likelihood function to find the optimize value for $B$ and $c$

```{r}
aiv=c(0.00016,1.09)
options(warn=-1)
u=nlm(Q5.7,aiv)$estimate
u
```
aiv vector is the initial value for nlm command. 
The option command is to ignore the warning (for presentation purpose only). The nlm command will find the value B,c for vector u(B,c) so that the negative likelihood (negative minimum = positive maximum) function Q5.7 is smallest.  
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of force of mortality for each age.
```{r}
ux=u[1]*u[2]^ax
data.frame("Age"=ax,"Force of mortality"=ux)
```

## Question 5.8
As given, we have:  
$ln(\dot{q}_x/(1-\dot{q}_x)=a+bx$  
First, we input the data.
```{r}
bx=c(30:49)
bEx=c(70000,66672,68375,65420,61779,66091,68514,69560,65000,66279,67300,65368,65391,62917,66537,62302,62145,63856,61097,61110)
bdx=c(39,43,34,31,23,50,48,43,48,47,62,63,84,86,120,121,122,162,151,184)
bcrude=bdx/bEx
lnq=log(bcrude/(1-bcrude))
```

Then, we create a weighted least squares function to minimize.  
$\sum_{x=30}^{49} E_x[ln(\hat{q_x}/1-\hat{q_x})-ln(\dot{q_x}/1-\dot{q_x})]^2$  
$\sum_{x=30}^{49} E_x[ln(\hat{q_x}/1-\hat{q_x})-a-bx]^2$  
In which: $\hat{q}=d_x/E_x$
```{r}
Q5.8=function(z){
  c=z[1]+z[2]*bx
  sum(bEx*(lnq-c)^2)
}
```

Now we will use the command nlm to minimize the weighted least square function to find the optimize value for $a$ and $b$
```{r}
biv=c(-10,0.09)
z=nlm(Q5.8,biv)$estimate
z
```
biv vector is the initial value for nlm command.  
The nlm command will find the value a,b for vector z(a,b) so that the function weighted least square Q5.8 will be minimize.
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of mortality rate for each age.

We have
$ln(\dot{q}_x/(1-\dot{q}_x))=a+bx$  
Therefore,
$\dot{q}_x=e^{a+bx}/(1+e^{a+bx})$
```{r}
e=exp(z[1]+z[2]*bx)
bqx=e/(e+1)
data.frame("Age"=bx,"Mortality rate"=bqx)
```
## Question 5.9
As given, we have:  
$\dot{q}_x=a+bq_x^S$  
First, we input the data.
```{r}
cx=c(47:67)
cEx=c(166,187,218,243,276,302,347,390,430,494,558,628,701,813,917,1040,1182,1299,1432,1596,1752)
cdx=c(2,2,4,6,2,4,7,3,9,9,8,11,14,18,18,24,30,43,41,54,64)
cqx=cdx/cEx
cqsx=c(0.00505,0.00570,0.00644,0.00728,0.00826,0.00930,0.01051,0.01184,0.01331,0.01492,0.01668,0.01859,0.02065,0.02287,0.02525,0.02778,0.03049,0.03339,0.03648,0.03978,0.04332)

```
Then, we create a weighted least squares function to minimize.  
$\sum_{x=47}^{67} E_x(\hat{q_x}-\dot{q_x})^2$  
$=\sum_{x=47}^{67} E_x(\hat{q_x}-a-bq_x^S)^2$  
In which: $\hat{q}=d_x/E_x$
```{r}
Q5.9=function(r){
  qs=r[1]+r[2]*cqsx
  sum(cEx*(cqx-qs)^2)
}
```
Now we will use the command nlm to minimize the weighted least square function to find the optimize value for $a$ and $b$
```{r}
civ=c(0.007,0.6)
r=nlm(Q5.9,civ)$estimate
r
```

civ vector is the initial value for nlm command.  
The nlm command will find the value a,b for vector z(a,b) so that the function weighted least square Q5.9 will be minimize.
The $estimate part is to extract the estimated values
Now we caculate the graduated estimates of mortality rate for each age.

We have
$\dot{q}_x=a+bq_x^S$  
```{r}
qs=r[1]+r[2]*cqsx
data.frame("Age"=cx,"Mortality rate"=qs)
```

