---
title: "Assignment 1 ACST886"
author: "Vu Nguyen"
date: "21/08/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the link to my Assignment on Github:  
https://github.com/vunguyen1502/ACST886/tree/master/Assignment%201  
https://github.com/vunguyen1502/ACST886  
Either link would do

# Question 1

I apply the annuity 
$R\times\sum_{n = 1}^{30} v^n=600,000$  
In which:  
$R$ is the annual payment  
$v$ is the discount factor with $v=(1+i)^{-1}$
  
```{r}
pmt=function(R) {
  i=0.035
  v=1/(1+i)
  n=c(1:30)
  a=sum(v^n)
  R*a-600000
}
```
The code create a function simulating the equation: 
$R\times\sum_{n = 1}^{30} v^n-600,000=0$  

Now we use uniroot command to find the solution for the equation.
```{r}
Q1=uniroot(pmt,interval = c(1,9999999))$root
Q1
```
The part $root extract the solution from the result table.
Testing the result:
```{r}
pmt(Q1)
```
This is the correct result.

# Question 2
## Part a)
Since Scott take on the same loan as Bill in Question 1 with the same term and interest rate, by not including the interest offset amount the initial payment of Scott and Bill should be the same.
```{r}
Q1
```
## Part b)
Since Scott have an offset account which pay annual interest at 3.5%, this means the annual repayment of Scott would be $R+100,000\times3.5\%=R+3,500$. With a greater annual repayment and a 100,000 payoff, the payment term should be less than 30 years. Prove:
```{r}
nper=function(n) {
  i=0.035
  v=1/(1+i)
    a=(1-v^n)/i
  (Q1+3500)*a-600000+100000*v^n
}
```
Since we expecting a non-integer anwser, I switch to an annuity formulae.
The code create a function simulating the equation: 
$(R+3,500)\times\frac{1-v^n}{i}+100,000v^n-600,000=0$  
Note that:  
$=\frac{1-v^n}{i}=\sum_{j = 1}^{n}v^j$  
$100,000\times v^n$ is the discounted value of the last payoff $100,000. 

```{r}
Q2a=uniroot(nper,interval = c(1,9999999))$root
Q2a
ceiling(Q2a)
```
The ceiling command round up the answer since the term can only be an interger. As we can see it only takes 23 years for Scott to payoff the loan  

Testing the result:  
```{r}
nper(Q2a)
```
The result is correct(Testing show value close to 0 due to non-integer term)

Since the calculated payment term is a non-integer, the last payment of Scott should be less than the initial payment. Prove:

```{r}
lpmt=function(z) {
  i=0.035
  v=1/(1+i)
    a=(1-v^23)/i
  (600000-(Q1+3500)*a)*(1+i)^23-z
}
```

I use the restrospective method to calculate the loan outstanding right before the last payment.
The code create a function simulating the equation:  
$[[600,000-(R+3,500)]\times\frac{1-v^{25}}{i}]\times(1+i)^{26}-z=0$  
In which:  
$z$ is the loan outstanding right before the offset account payoff.

```{r}
z=uniroot(lpmt,interval = c(1,9999999))$root
z
```

Scott last repayment will be:

```{r}
Q2b=z+Q1-100000
Q2b
```

The result shows that the last payment is less than the initial payment calculated in part a.
In conclusion, with the offset account in effect, the payment term of Scott would be 23 years with the last repayment is $11,484.89.

# Question 3
## Part a)
Theresa receives $1,000 at the end of each year for 30 years. The total present value of the commission would be:  
$1,000\times\frac{1-v^{30}}{i}$  
In which:  
$i=3.5\%$ and $v=(1+i)^{-1}$
```{r}
i=0.035
v=1/(1+i)
Q3a=1000*(1-v^30)/i
Q3a
```

## Part b)
We have:
$L=R\times\sum_{j = 1}^{n} (1+i)^{-j}$  
In which:  
$L$ is the total loan amount  
$n$ is the loan term  
$i$ is the rate  
$R$ is the periodical repayment amount  
As you can see, as the loan amount $L$ increase and term $n$ and $R$ stay the same, the effective interest $i$ will decrease  
$~$  
$~$  


## Part c)
To calculate the effective rate, we will create a function simulate the equation on part b).
```{r}
rate=function(i) {
  v=1/(1+i)
  n=c(1:30)
  a=(1-v^30)/i
  Q1*a-600000-Q3a
}
```
The code create a function simulating the equation:
$R\times\sum_{n = 1}^{n} (1+i)^{-j}-L=0$  
In which:  
$L$ is the sum of the loan amount and the total PV of Theresa commission at 19 March 2019.
```{r}
Q3c=uniroot(rate,interval = c(0.000000001,9999999))$root
Q3c
```
The uniroot command  will find a interest rate $i$ which with all the given input ($L, n, R$), the equation simulated in function $rate$ will result in 0. This will be our effective rate (3.26%).  

## Part d)
If Bill was offer the interest rate in part b), his annual repayment (assuming the loan term and amount stay the same) would be:
```{r}
i=Q3c
v=1/(1+i)
n=c(1:30)
Q3d=600000/sum(v^n)
Q3d
```
The new annual repayment is $31,652.49
```{r}
(Q1-Q3d)/Q1
```
Bill's annual repayment decreases by 2.97%
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  

# Question 4
Input data  

```{r}
LifeA <- as.Date(c("1965-05-16", "1985-07-21", "1994-06-22"))
LifeB <- as.Date(c("1922-03-01", "1955-09-23", "1993-07-21"))
LifeC <- as.Date(c("1942-09-21", "1967-06-02", "1996-04-29"))
```
## Part a) Age next birthday at death

```{r}
LifePartA=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-montha)*32+(dayc-daya))>=0){
  yearc-yeara+1
} else {yearc-yeara}
}
```
- The first line is to create a function to calculate age next birthday.  
- The next 6 lines are to extract the date, month, year value for calculate.  
- The 8th line indicate the condition whether the Life die before or after birthday within the year of dead. The next 2 lines give coresponding results.  

The code to check the result is as below.
```{r}
data.frame("Age of life A"=LifePartA(LifeA), "Age of life B"=LifePartA(LifeB), 
           "Age of life C"=LifePartA(LifeC))
```


## Part b) Age last birthday an 1 January preceding death
```{r}
LifePartB=function(Life){
montha=as.numeric(substr(Life[1],6,7))
daya=as.numeric(substr(Life[1],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (montha!=1&&daya!=1){
  yearc-yeara-1
} else {yearc-yeara}
}
```
- I extract the value just like in part a)  
- For the condition, if the birthday of the life chosen is not in 01 January, we simple take the difference between death year and birth year minus 1. Otherwise, we won't minus one.  
  
The code to check the result is as below.

```{r}
data.frame("Age of life A"=LifePartB(LifeA), "Age of life B"=LifePartB(LifeB), 
           "Age of life C"=LifePartB(LifeC))

```

## Part c) Age at the birthday in the policy year of death
```{r}
LifePartC=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthb=as.numeric(substr(Life[2],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayb=as.numeric(substr(Life[2],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearb=as.numeric(substr(Life[2],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-monthb)*32+(dayc-dayb))>0){
  z=yearc
} else {z=yearc-1}
if (((monthb-montha)*32+(dayb-daya))>0){
  z+1-yeara
} else {z-yeara}
}

```
- The first 10 line is to create the function to calculate age and extract the date value like part a and part b.  
- Line 11 to line 13 state the condition and extract the policy year at death (it will take value 1993 if policy year at death is 7/1993-7/1994).  
- If the death date is before the policy date (in the same calendar year), then the policy year at death will be the death year minus 1. Otherwise don't minus 1.   
- Line 14 to line 16 state the condition and calculate the exact age at birthday at policy year of death.  
- If the birthday is before the policy date (in the same calendar year), then the age will be the policy year at death minus birthday year plus 1. Otherwise, don't plus 1.  
  
The code to check the result is as below.
```{r}
data.frame("Age of life A"=LifePartC(LifeA),"Age of life B"=LifePartC(LifeB), 
           "Age of life C"=LifePartC(LifeC))
```

