---
title: "Question 2 Assignment 1 ACST886"
author: "Vu Nguyen"
date: "22/08/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2
## Part a)
Since Scott take on the same loan as Bill in Question 1 with the same term and interest rate, by not including the interest offset amount the initial payment of Scott and Bill should be the same.
```{r}
Q1=32622.8
Q1
```
## Part b)
Since Scott have an offset account which pay annual interest at 3.5%, this means the annual repayment of Scott would be $R+100,000\times3.5\%=R+3,500$. With a greater annual repayment and a 100,000 payoff, the payment term should be less than 30 years. Prove:
```{r}
nper=function(n) {
  i=0.035
  v=1/(1+i)
    a=(1-v^n)/i
  (Q1+3500)*a-600000+100000*v^n
}
```
Since we expecting a non-integer anwser, I switch to an annuity formulae.
The code create a function simulating the equation: 
$(R+3,500)\times\frac{1-v^n}{i}+100,000v^n-600,000=0$  
Note that:  
$=\frac{1-v^n}{i}=\sum_{j = 1}^{n}v^j$  
$100,000\times v^n$ is the discounted value of the last payoff $100,000. 

```{r}
Q2a=uniroot(nper,interval = c(1,9999999))$root
Q2a
ceiling(Q2a)
```
The ceiling command round up the answer since the term can only be an interger. As we can see it only takes 23 years for Scott to payoff the loan  

Testing the result:  
```{r}
nper(Q2a)
```
The result is correct(Testing show value close to 0 due to non-integer term)

Since the calculated payment term is a non-integer, the last payment of Scott should be less than the initial payment. Prove:

```{r}
lpmt=function(z) {
  i=0.035
  v=1/(1+i)
    a=(1-v^23)/i
  (600000-(Q1+3500)*a)*(1+i)^23-z
}
```

I use the restrospective method to calculate the loan outstanding right before the last payment.
The code create a function simulating the equation:  
$[[600,000-(R+3,500)]\times\frac{1-v^{25}}{i}]\times(1+i)^{26}-z=0$  
In which:  
$z$ is the loan outstanding right before the offset account payoff.

```{r}
z=uniroot(lpmt,interval = c(1,9999999))$root
z
```

Scott last repayment will be:

```{r}
Q2b=z+Q1-100000
Q2b
```

The result shows that the last payment is less than the initial payment calculated in part a.
In conclusion, with the offset account in effect, the payment term of Scott would be 23 years with the last repayment is $11,484.89