---
title: "Question 3 Assignment 1 ACST886"
author: "Vu Nguyen"
date: "22/08/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 3
## Part a)
Theresa receives $1,000 at the end of each year for 30 years. The total present value of the commission would be:  
$1,000\times\frac{1-v^{30}}{i}$  
In which:  
$i=3.5\%$ and $v=(1+i)^{-1}$
```{r}
i=0.035
v=1/(1+i)
Q3a=1000*(1-v^30)/i
Q3a
```

## Part b)
We have:
$L=R\times\sum_{j = 1}^{n} (1+i)^{-j}$  
In which:  
$L$ is the total loan amount  
$n$ is the loan term  
$i$ is the rate  
$R$ is the periodical repayment amount  
As you can see, as the loan amount $L$ increase and term $n$ and $R$ stay the same, the effective interest $i$ will decrease

## Part c)
To calculate the effective rate, we will create a function simulate the equation on part b).
```{r}
Q1=32622.8
rate=function(i) {
  v=1/(1+i)
  n=c(1:30)
  a=(1-v^30)/i
  Q1*a-600000-Q3a
}
```
The code create a function simulating the equation:
$R\times\sum_{n = 1}^{n} (1+i)^{-j}-L=0$  
In which:  
$L$ is the sum of the loan amount and the total PV of Theresa commission at 19 March 2019.
```{r}
Q3c=uniroot(rate,interval = c(0.000000001,9999999))$root
Q3c
```
The uniroot command  will find a interest rate $i$ which with all the given input ($L, n, R$), the equation simulated in function $rate$ will result in 0. This will be our effective rate (3.26%).  

## Part d)
If Bill was offer the interest rate in part b), his annual repayment (assuming the loan term and amount stay the same) would be:
```{r}
i=Q3c
v=1/(1+i)
n=c(1:30)
Q3d=600000/sum(v^n)
Q3d
```
The new annual repayment is $31,652.49
```{r}
(Q1-Q3d)/Q1
```
Bill's annual repayment decreases by 2.97%
