---
title: "Question 2 Assignment 3_ACST886"
author: "Vu Nguyen"
date: "02/10/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My code is written using R Markdown
BE NOTICED:  
Question 3 part a) is written at the end of Question 2 file  

This is the link to my Assignment on Github:  
https://github.com/vunguyen1502/ACST886/tree/master/Assignment%203  
https://github.com/vunguyen1502/ACST886  
Either link would do.

# Question 3 part a)
Refer to the end of Question 2 file.  
The yield is 47.38256%.  

# Question 3 part b)
This is the same as question 2 except we have to adjust the expenses and add an insurance charge to tha table.  

## Generate double decrement table
Here we need vectors of  

1. Age  
2. Dependent rate  

### First we create a vector of age
```{r}
x=c(30:39)
x
```
### Death rate (independent)
```{r}
iqdx=0.00001*c(43,42,41,40,41,42,43,45,49,53)
iqdx
```
### Withdrawal rate (independent)
```{r}
iqwx=c(0.25,0.2,0.175,0.15,0.15,0.15,0.15,0.15,0.15,1)
iqwx
```

### Death rate (dependent)
The formula is $(aq)_x^d=q_x^d(1-0.5q_x^w)$
```{r}
qdx=iqdx*(1-0.5*iqwx)
qdx
```

### Withdrawal rate (dependent)
The formula is $(aq)_x^w=q_x^w(1-0.5q_x^d)$
```{r}
qwx=iqwx*(1-0.5*iqdx)
qwx
```

### Double decrement table
Now we create the double decrement table with the dependent rate and radix 100,000.
```{r}
alx=c()
add=c()
adw=c()
alx[1]=100000
for (i in c(1:10)){
  add[i]=alx[i]*qdx[i]
  adw[i]=alx[i]*qwx[i]
  alx[i+1]=alx[i]-add[i]-adw[i]
}
```
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  
$~$  

Create a data frame for visualizing
```{r}
data.frame("Age"=x,"qd"=iqdx,"qw"=iqwx,"aqd"=qdx,"aqw"=qwx,"al"=alx[1:10],"add"=add,"adw"=adw)
```

### Insurance charge
The formula for insurance charge is:  
$S\times 1.2q_x^d$ with $S$ is the sum insured
```{r}
Ins.chr=20000*iqdx*1.2
Ins.chr
```

## Unit fund in-force cashflow  
Here we need vectors of  

1. Invested premium  
2. Value of unit  

### Invested premium
We can calculated invested premium using the formula:    
$IP_t=P_ta_t(1-b)-insurance~charge$. In which:  
$IP_t$ is the invested premium of year t.  
$P_t$ is the premium of year t.  
$1-a$ is the premium base fee.  
$b$ is the buy sell margin.  

```{r}
Invest.Prem=rep(5000*0.99*0.995,10)-Ins.chr
Invest.Prem[1]=5000*0.55*0.995-Ins.chr[1]
Invest.Prem
```

### Value of unit
We can calculated value of unit using the formula: 
$VU_t=(VU_{t-1}+IP_t)(1+i)(1-m)$. In Which:  
$VU_t$ is the value of unit at time $t$.  
$m$ is the asset fee.  
$i$ is the earning rate of unit fund.  
When $t=1$, $VU_t=IP_0(1+i)(1-m)$
```{r}
Value.Unit=c()
Value.Unit[1]=Invest.Prem[1]*1.1*0.997
for (i in c(2:10)){
  Value.Unit[i]=(Value.Unit[i-1]+Invest.Prem[i])*1.1*0.997
}
Value.Unit
```
### Unit fund in-force cashflow table
```{r}
data.frame("Year"=c(1:10), "Invested premium.BoY"=Invest.Prem,"Value of Unit.EoY"=Value.Unit)
```

## Unit fund expected amounts per initial policy
Here we need vectors of  

1. Expected invested premium  
2. Unit fund interest  
3. Expected value of unit  
4. Transfer to non-unit fund
$~$  
$~$  
$~$  

### Invested premium
```{r}
Ex_Invest.Prem=Invest.Prem*alx[1:10]/alx[1]
Ex_Invest.Prem
```

### Value of units
```{r}
Ex_Value.Unit=Value.Unit*alx[2:11]/alx[1]
Ex_Value.Unit
```

### Interest
```{r}
UF.Int=c()
UF.Int[1]=Ex_Invest.Prem[1]*0.1
UF.Int[2:10]=(Ex_Invest.Prem[2:10]+Ex_Value.Unit[1:9])*0.1
UF.Int
```

### Transfer to Non-unit fund
```{r}
UF_Transfer=c()
UF_Transfer=Ex_Invest.Prem[1]+UF.Int[1]-Ex_Value.Unit[1]
UF_Transfer[2:10]=Ex_Invest.Prem[2:10]+UF.Int[2:10]-Ex_Value.Unit[2:10]+Ex_Value.Unit[1:9]
UF_Transfer
```

### Unit fund expected amounts per initial policy
We create a unit fund table to easily visualize the result.  
```{r}
data.frame("Year"=c(1:10),"Invested premium"=Invest.Prem,"Value of Units"=Value.Unit,
           "Ex Invested premium"=Ex_Invest.Prem,"Interest"=UF.Int,
           "Ex Value of Units"=Ex_Value.Unit,"Transfer to Non-unit fund"=UF_Transfer)
```

## Non-unit fund in-force cashflow
Here we need vectors of  

1. Premium base fees  
2. Buy-sell margin  
3. Expenses  
4. Surrender benefit  
5. Death benefit  
6. Insurance charge  

### Premium base fee
The formula for premium base fee is $P_t(1-a_t)$ with $(1-a_t)$ is the premium based fee for year $t$.
```{r}
Prem.fee=rep(5000*0.01,10)
Prem.fee[1]=5000*0.45
Prem.fee
```
$~$  

### Buy-sell margin
The formula for the buy-sell margin is $bP_ta_t$ with b is the buy/sell margin.  
```{r}
BS.Margin=rep(5000*0.99*0.005,10)
BS.Margin[1]=5000*(1-0.45)*0.005
BS.Margin
```

### Expenses
```{r}
Exp=75*1.25^c(0:9)
Exp[1]=75+5000*0.45
Exp
```
### Surrender benefit
This is the same as sell value of unit
```{r}
Surr.bft=Value.Unit
Surr.bft
```

### Death benefit
This is the same as sell value of unit plus the insured amount ($20,000)
```{r}
Death.bft=Value.Unit+20000
Death.bft
```

### Insurance charge
As calculated above
```{r}
Ins.chr
```
$~$  

### Non-unit fund in-force cashflow table  
```{r}
data.frame("Year"=c(1:10),"Premium-based fees"=Prem.fee,
           "Buy-sell Margins"=BS.Margin,
           "Expenses"=Exp,"Surr.bft"=Surr.bft,
           "Death.bft"=Death.bft,"Ins charge"=Ins.chr)
```
## Non-unit fund expected amounts per initial policy
Here we need vectors of  

1. Expected premium base fees  
2. Expected buy /sell margin  
3. Expected expenses  
4. Non-unit fund interest  
5. Expected death benefit  
6. Expected surrender benefit  
7. Transfer from unit fund  
8. Expected insurance charge  
9. Transfer

### Expected premium based fee
```{r}
Ex_Prem.fee=Prem.fee*alx[1:10]/alx[1]
Ex_Prem.fee
```

### Expected buy-sell margin
```{r}
Ex_BS.margin=BS.Margin*alx[1:10]/alx[1]
Ex_BS.margin
```

### Expected expenses
```{r}
Ex_Exp=Exp*alx[1:10]/alx[1]
Ex_Exp
```

### Expected insurance charge
```{r}
Ex_Ins.chr=Ins.chr*alx[1:10]/alx[1]
Ex_Ins.chr
```

### Interest
We add the expected insurance charge to the original interest formula
```{r}
NUF_Int=(Ex_Prem.fee-Ex_Exp+Ex_BS.margin+Ex_Ins.chr)*0.08
NUF_Int
```
$~$  

### Death benefit
```{r}
Ex_D.benefit=Death.bft*add/alx[1]
Ex_D.benefit
```

### Surrender benefit
```{r}
Ex_S.benefit=Surr.bft*adw/alx[1]
Ex_S.benefit
```

### Profit signature (Transfer)
The profit signature is:
```{r}
NUF_Transfer=Ex_Prem.fee+Ex_BS.margin-Ex_Exp+NUF_Int+UF_Transfer-Ex_D.benefit-Ex_S.benefit+Ex_Ins.chr
NUF_Transfer
```
### Non-unit fund expected amounts per initial policy table
```{r}
data.frame("Year"=c(1:10),"Premium-based fees"=Ex_Prem.fee,
           "Buy-sell Margins"=Ex_BS.margin,
           "Expenses"=Ex_Exp,"Interest"=NUF_Int,
           "Trns from Unit fund"=UF_Transfer,
           "Death.bft"=Ex_D.benefit,
           "Surr.bft"=Ex_S.benefit,
           "Ins charge"=Ex_Ins.chr,"Transfer"=NUF_Transfer)
```
### Calculate EPV and Yield
Now we calculate the EPV and yield
The EPV of transfers is:
```{r}
EPV=sum(NUF_Transfer*((1/1.125)^c(1:10)))
EPV
```

The yield is:
```{r}
function_epv=function(i){
  sum(NUF_Transfer*((1+i)^(-c(1:10))))
}
irr=uniroot(function_epv,c(-99,1))
irr$root
```
Testing yield:  
```{r}
round(function_epv(irr$root),2)
```

Result:  
$EPV=-114.6149$  
$Yield=-192.407\%$