---
title: "Question 2 Assignment 3_ACST886"
author: "Vu Nguyen"
date: "02/10/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My code is written using R Markdown  
BE NOTICED:  
Question 3 part a) is written at the end of this file

This is the link to my Assignment on Github:  
https://github.com/vunguyen1502/ACST886/tree/master/Assignment%203  
https://github.com/vunguyen1502/ACST886  
Either link would do.

# Question 2 part a)
## Generate double decrement table
Here we need vectors of

1. Age
2. Dependent rate  

### First we create a vector of age
```{r}
x=c(30:39)
x
```
### Death rate (independent)
```{r}
iqdx=0.00001*c(43,42,41,40,41,42,43,45,49,53)
iqdx
```
### Withdrawal rate (independent)
```{r}
iqwx=c(0.25,0.2,0.175,0.15,0.15,0.15,0.15,0.15,0.15,1)
iqwx
```

### Death rate (dependent)
The formula is $(aq)_x^d=q_x^d(1-0.5q_x^w)$
```{r}
qdx=iqdx*(1-0.5*iqwx)
qdx
```

### Withdrawal rate (dependent)
The formula is $(aq)_x^w=q_x^w(1-0.5q_x^d)$
```{r}
qwx=iqwx*(1-0.5*iqdx)
qwx
```

### Double decrement table
Now we create the double decrement table with the dependent rate and radix 100,000.
```{r}
alx=c()
add=c()
adw=c()
alx[1]=100000
for (i in c(1:10)){
  add[i]=alx[i]*qdx[i]
  adw[i]=alx[i]*qwx[i]
  alx[i+1]=alx[i]-add[i]-adw[i]
}
```

Create a data frame for visualizing
```{r}
data.frame("Age"=x,"Independent Death rate"=iqdx,"Independent Withdralwal rate rate"=iqwx,
           "aq.d"=qdx,"aq.w"=qwx,"al"=alx[1:10],"ad.d"=add,"ad.w"=adw)
```

## Unit fund in-force cashflow  
Here we need vectors of  

1. Invested premium  
2. Value of unit  

### Invested premium
We can calculated invested premium using the formula:    
$IP_t=P_ta_t(1-b)$. In which:  
$IP_t$ is the invested premium of year t.  
$P_t$ is the premium of year t.  
$1-a$ is the premium base fee.  
$b$ is the buy sell margin.  

```{r}
Invest.Prem=rep(5000*0.99*0.995,10)
Invest.Prem[1]=5000*0.55*0.995
Invest.Prem
```

### Value of unit
We can calculated value of unit using the formula: 
$VU_t=(VU_{t-1}+IP_t)(1+i)(1-m)$. In Which:  
$VU_t$ is the value of unit at time $t$.  
$m$ is the asset fee.  
$i$ is the earning rate of unit fund.  
When $t=1$, $VU_t=IP_0(1+i)(1-m)$
```{r}
Value.Unit=c()
Value.Unit[1]=Invest.Prem[1]*1.1*0.997
for (i in c(2:10)){
  Value.Unit[i]=(Value.Unit[i-1]+Invest.Prem[i])*1.1*0.997
}
Value.Unit
```
### Unit fund in-force cashflow table
```{r}
data.frame("Year"=c(1:10), "Invested premium"=Invest.Prem,"Value of Unit"=Value.Unit)
```


## Unit fund expected amounts per initial policy
Here we need vectors of  

1. Expected invested premium  
2. Unit fund interest  
3. Expected value of unit  
4. Transfer to non-unit fund  

### Invested premium
```{r}
Ex_Invest.Prem=Invest.Prem*alx[1:10]/alx[1]
Ex_Invest.Prem
```

### Value of units
```{r}
Ex_Value.Unit=Value.Unit*alx[2:11]/alx[1]
Ex_Value.Unit
```
$~$  

### Interest
```{r}
UF.Int=c()
UF.Int[1]=Ex_Invest.Prem[1]*0.1
UF.Int[2:10]=(Ex_Invest.Prem[2:10]+Ex_Value.Unit[1:9])*0.1
UF.Int
```

### Transfer to Non-unit fund
```{r}
UF_Transfer=c()
UF_Transfer=Ex_Invest.Prem[1]+UF.Int[1]-Ex_Value.Unit[1]
UF_Transfer[2:10]=Ex_Invest.Prem[2:10]+UF.Int[2:10]-Ex_Value.Unit[2:10]+Ex_Value.Unit[1:9]
UF_Transfer
```
### Unit fund expected amounts per initial policy

We create a unit fund table to easily visualize the result.  
```{r}
data.frame("Year"=c(1:10),"Invested premium"=Invest.Prem,"Value of Units"=Value.Unit,
           "Ex Invested premium"=Ex_Invest.Prem,"Interest"=UF.Int,
           "Ex Value of Units"=Ex_Value.Unit,"Transfer to Non-unit fund"=UF_Transfer)
```

## Non-unit fund in-force cashflow
Here we need vectors of  

1. Premium base fees  
2. Buy-sell margin  
3. Expenses  
4. Death/Surrender benefit  

### Premium base fee
The formula for premium base fee is $P_t(1-a_t)$ with $(1-a_t)$ is the premium based fee for year $t$.
```{r}
Prem.fee=rep(5000*0.01,10)
Prem.fee[1]=5000*0.45
Prem.fee
```

### Buy-sell margin
The formula for the buy-sell margin is $bP_ta_t$ with b is the buy/sell margin.  
```{r}
BS.Margin=rep(5000*0.99*0.005,10)
BS.Margin[1]=5000*(1-0.45)*0.005
BS.Margin
```

### Expenses
```{r}
t.series=c(0:9)
Exp=58*1.2^t.series
Exp[1]=58+5000*0.45
Exp
```
$~$  
$~$  
$~$  
$~$  

### Death/surrender benefit
This is the same as sell value of unit
```{r}
DS.benefit=Value.Unit
DS.benefit
```

### Non-unit fund in-force cashflow table  
```{r}
data.frame("Year"=c(1:10),"Premium-based fees"=Prem.fee,
           "Buy-sell Margins"=BS.Margin,
           "Expenses"=Exp,"Death/surr.bft"=DS.benefit)
```

## Non-unit fund expected amounts per initial policy
Here we need vectors of  

1. Expected premium base fees  
2. Expected buy /sell margin  
3. Expected expenses  
4. Non-unit fund interest  
5. Expected death benefit  
6. Expected surrender benefit  
7. Transfer from unit fund  
8. Transfer

### Expected premium based fee
```{r}
Ex_Prem.fee=Prem.fee*alx[1:10]/alx[1]
Ex_Prem.fee
```

### Expected buy-sell margin
```{r}
Ex_BS.margin=BS.Margin*alx[1:10]/alx[1]
Ex_BS.margin
```

### Expected expenses
```{r}
Ex_Exp=Exp*alx[1:10]/alx[1]
Ex_Exp
```

### Interest
```{r}
NUF_Int=(Ex_Prem.fee-Ex_Exp+Ex_BS.margin)*0.08
NUF_Int
```

### Death benefit
```{r}
Ex_D.benefit=DS.benefit*add/alx[1]
Ex_D.benefit
```
$~$  
$~$  

### Surrender benefit
```{r}
Ex_S.benefit=DS.benefit*adw/alx[1]
Ex_S.benefit
```
### Transfer from unit fund
```{r}
UF_Transfer
```

### Profit signature (Transfer)
The profit signature is:
```{r}
NUF_Transfer=Ex_Prem.fee+Ex_BS.margin-Ex_Exp+NUF_Int+UF_Transfer-Ex_D.benefit-Ex_S.benefit
NUF_Transfer
```

### Non-unit fund expected amounts per initial policy table
```{r}
data.frame("Year"=c(1:10),"Premium-based fees.BoY"=Ex_Prem.fee,
           "Buy-sell Margins.BoY"=Ex_BS.margin,
           "Expenses.BoY"=Ex_Exp,
           "Interest.EoY"=NUF_Int,
           "Trns from Unit fund.EoY"=UF_Transfer,
           "Death.bft.EoY"=Ex_D.benefit,
           "Surr.bft.EoY"=Ex_S.benefit,
           "Transfer"=NUF_Transfer)
```


### Determine the profit signature
```{r}
NUF_Transfer
```


### Calculate the expected NPV
Now we calculate the expected NPV.  
The expected NPV is:
```{r}
NPV=sum(NUF_Transfer*((1/1.125)^c(1:10)))
NPV
```

# Question 2 part b)  

First, we will create a function to the 10th value of the transfer equal zero by inputing the 9th value of the policy value vector.  
Then we use the uniroot command to find the 9th value of the policy vector so that the 10th transfer equal zero.
```{r}
value=function(z){
# Set the 9th policy as a variable
Value.Unit[9]=z

# Revaluate affected vector
## Expected policy value
Ex_Value.Unit=Value.Unit*alx[2:11]/alx[1]

## Interest (Unit fund)
UF.Int=c()
UF.Int[1]=Ex_Invest.Prem[1]*0.1
UF.Int[2:10]=(Ex_Invest.Prem[2:10]+Ex_Value.Unit[1:9])*0.1

## Transfer to non-unit fund
UF_Transfer=c()
UF_Transfer=Ex_Invest.Prem[1]+UF.Int[1]-Ex_Value.Unit[1]
UF_Transfer[2:10]=Ex_Invest.Prem[2:10]+UF.Int[2:10]-Ex_Value.Unit[2:10]+Ex_Value.Unit[1:9]

## Death/surrender benefit
DS.benefit=Value.Unit

## Expected death benefit
Ex_D.benefit=DS.benefit*add/alx[1]

## Expected surrender benefit
Ex_S.benefit=DS.benefit*adw/alx[1]

## Transfer
NUF_Transfer=Ex_Prem.fee+Ex_BS.margin-Ex_Exp+NUF_Int+UF_Transfer-Ex_D.benefit-Ex_S.benefit

## 10th transfer
NUF_Transfer[10]
}
```
  
Now we use the uniroot command to find the 9th value of the policy vector so that the 10th transfer equal zero.
```{r}
uniroot(value,c(0,999999))
Policy.value=uniroot(value,c(0,999999))$root
Policy.value
```
Therefore, the 9th value we need to find is 67,319.02  
Testing:  
```{r}
round(value(Policy.value),11)
```

# Question 2 part c)
We can adjust the function in part b) to give us the new expected
NPV of transfer.
```{r}
new.npv=function(z){
# Set the 9th policy as a variable
Value.Unit[9]=z

# Revaluate affected vector
## Expected policy value
Ex_Value.Unit=Value.Unit*alx[2:11]/alx[1]

## Interest (Unit fund)
UF.Int=c()
UF.Int[1]=Ex_Invest.Prem[1]*0.1
UF.Int[2:10]=(Ex_Invest.Prem[2:10]+Ex_Value.Unit[1:9])*0.1

## Transfer to non-unit fund
UF_Transfer=c()
UF_Transfer=Ex_Invest.Prem[1]+UF.Int[1]-Ex_Value.Unit[1]
UF_Transfer[2:10]=Ex_Invest.Prem[2:10]+UF.Int[2:10]-Ex_Value.Unit[2:10]+Ex_Value.Unit[1:9]

## Death/surrender benefit
DS.benefit=Value.Unit

## Expected death benefit
Ex_D.benefit=DS.benefit*add/alx[1]

## Expected surrender benefit
Ex_S.benefit=DS.benefit*adw/alx[1]

## Transfer
NUF_Transfer=Ex_Prem.fee+Ex_BS.margin-Ex_Exp+NUF_Int+UF_Transfer-Ex_D.benefit-Ex_S.benefit

## Expected NPV
sum(NUF_Transfer*1.125^(-c(1:10)))
}
n.npv=new.npv(Policy.value)
n.npv
```
Now for comparision:
```{r}
n.npv<NPV
```

The result show that the expected present value of transfer at 12.5% p.a has decreased.  
This is because, money has been held back in the Unit Fund at time 9 (rather than being released to shareholders) to give a zero transfer at time 10. Since these additional funds earn only 10% p.a. interest in the Unit Fund, whereas the discount rate is 12.50%, overall the business is now (slightly) less profitable.

# Question 3 part a)
We create a function to calculate the EPV of transfer based on the input discount rate.  
```{r}
airr=function(i){
  sum(NUF_Transfer*((1+i)^-c(1:10)))
}
```
Now we use the uniroot command to find the yield.  
```{r}
uniroot(airr,c(0.01,0.99))
yield=uniroot(airr,c(0.01,0.99))$root
yield
```
The yield is 47.38256%.
