---
title: "Question 4 Assignment 1 ACST886"
author: "Vu Nguyen"
date: "22/08/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 4
Input data  
```{r}
LifeA <- as.Date(c("1965-05-16", "1985-07-21", "1994-06-22"))
LifeB <- as.Date(c("1922-03-01", "1955-09-23", "1993-07-21"))
LifeC <- as.Date(c("1942-09-21", "1967-06-02", "1996-04-29"))
```
## Part a) Age next birthday at death

```{r}
LifePartA=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-montha)*32+(dayc-daya))>=0){
  yearc-yeara+1
} else {yearc-yeara}
}
```
- The first line is to create a function to calculate age next birthday.  
- The next 6 lines are to extract the date, month, year value for calculate.  
- The 8th line indicate the condition whether the Life die before or after birthday within the year of dead. The next 2 lines give coresponding results.  

The code to check the result is as below.
```{r}
data.frame("Age of life A"=LifePartA(LifeA), "Age of life B"=LifePartA(LifeB), 
           "Age of life C"=LifePartA(LifeC))
```


## Part b) Age last birthday an 1 January preceding death
```{r}
LifePartB=function(Life){
montha=as.numeric(substr(Life[1],6,7))
daya=as.numeric(substr(Life[1],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (montha!=1&&daya!=1){
  yearc-yeara-1
} else {yearc-yeara}
}
```
- I extract the value just like in part a)  
- For the condition, if the birthday of the life chosen is not in 01 January, we simple take the difference between death year and birth year minus 1. Otherwise, we won't minus one.  
  
The code to check the result is as below.

```{r}
data.frame("Age of life A"=LifePartB(LifeA), "Age of life B"=LifePartB(LifeB), 
           "Age of life C"=LifePartB(LifeC))

```

## Part c) Age at the birthday in the policy year of death
```{r}
LifePartC=function(Life){
montha=as.numeric(substr(Life[1],6,7))
monthb=as.numeric(substr(Life[2],6,7))
monthc=as.numeric(substr(Life[3],6,7))
daya=as.numeric(substr(Life[1],9,10))
dayb=as.numeric(substr(Life[2],9,10))
dayc=as.numeric(substr(Life[3],9,10))
yeara=as.numeric(substr(Life[1],1,4))
yearb=as.numeric(substr(Life[2],1,4))
yearc=as.numeric(substr(Life[3],1,4))
if (((monthc-monthb)*32+(dayc-dayb))>0){
  z=yearc
} else {z=yearc-1}
if (((monthb-montha)*32+(dayb-daya))>0){
  z+1-yeara
} else {z-yeara}
}

```
- The first 10 line is to create the function to calculate age and extract the date value like part a and part b.  
- Line 11 to line 13 state the condition and extract the policy year at death (it will take value 1993 if policy year at death is 7/1993-7/1994).  
- If the death date is before the policy date (in the same calendar year), then the policy year at death will be the death year minus 1. Otherwise don't minus 1.   
- Line 14 to line 16 state the condition and calculate the exact age at birthday at policy year of death.  
- If the birthday is before the policy date (in the same calendar year), then the age will be the policy year at death minus birthday year plus 1. Otherwise, don't plus 1.  
  
The code to check the result is as below.
```{r}
data.frame("Age of life A"=LifePartC(LifeA),"Age of life B"=LifePartC(LifeB), 
           "Age of life C"=LifePartC(LifeC))
```

